---
title: "Argentina Covid Report"
author: "andino"
date: "7/20/2020"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(incidence)
library (projections)
library(epitrix)
library(EpiModel)
library(EpiEstim)
library(rmarkdown)
library(knitr)
library(distcrete)
library(Hmisc)
require(RCurl)
require(googlesheets4)
require(tidyr)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = '!h')
options(knitr.kable.NA = '')

# Load data from Google Sheet
data <- read.csv("https://raw.githubusercontent.com/andinocl/arg_covid/master/data/arg_covid.csv",header=TRUE,sep=",")
covid_frame <- data.frame(data)
covid_frame[,c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19)] <- sapply(covid_frame[,c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19)], as.numeric)
rows <- nrow(covid_frame)-1
covid_frame[,1] <- as.Date(covid_frame[,1],"%Y-%m-%d")

# Set up for loops
national <- c("Argentina","NewCases","NewCases7Day","light blue","blue")
CABA <- c("CABA","CABANewCases","CABANewCases7Day","green","dark green")
PBA <- c("Province of Buenos Aires","PBANewCases","PBANewCases7Day","pink","red")
locality = array(c(national,CABA,PBA),
                 dim=c(5,3)
                 )
colnames(locality) <- c("Argentina","CABA","PBA")
rownames(locality) <- c("title","cases","avg","light","dark")
x <- colnames(locality)
```

## New Cases

The following graphs show the overall epidemiological curves in the localities:

```{r}
# plot total cases
for(val in x) {
  print(val)
  x_val = covid_frame[1:rows,1]
  y_var = locality["cases",val]
  y_val = covid_frame[1:rows,y_var]
  plot(covid_frame[1:rows,1],
       covid_frame[1:rows,locality["cases",val]],
       type="n",
       xlab="Date", 
       ylab="Cases",
       main=paste("Daily new cases, ",locality["title",val])
  )
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  minor.tick(nx=2,ny=2)
  lines(covid_frame[1:rows,1],
        covid_frame[1:rows,locality["cases",val]],
        col=locality["light",val])
  lines(covid_frame[1:rows,1],
        covid_frame[1:rows,locality["avg",val]],
        col=locality["dark",val])
  legend(x="topleft",
         y=max(covid_frame[1:rows,locality["cases",val]],na.rm=T),
         legend=c("Daily Report","7-Day Rolling Average"),
         col=c(locality["light",val],locality["dark",val]),
         lty=1:2, cex=0.8)
}

```

## 14-day trend

Zooming in on the 14-day trend lines:

```{r}
# plot 14-day trend
for(val in x) {
  start_x = rows-14
  plot(covid_frame[start_x:rows,1],
       covid_frame[start_x:rows,locality["cases",val]],
       type="n",
       xlab="Date", 
       ylab="Cases",
       main=paste("Daily new cases, ",locality["title",val])
  )
  minor.tick(ny=2)
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  lines(covid_frame[start_x:rows,1],
        covid_frame[start_x:rows,locality["cases",val]],
        col=locality["light",val])
  lines(covid_frame[start_x:rows,1],
        covid_frame[start_x:rows,locality["avg",val]],
        col=locality["dark",val])
  legend(x="topleft",
         y=max(covid_frame[1:rows,locality["cases",val]],na.rm=T),
         legend=c("Daily Report","7-Day Rolling Average"),
         col=c(locality["light",val],locality["dark",val]),
         lty=1:2, cex=0.8)
}
```

## Log graphs

The following graphs are generated by:
$$
x = Number\ of\ Days\ since\ March\ 3\\
y = log(Number\ of\ New\ Cases\ this\ day)
$$
The regression line is drawn using the R "lm()" function over the x values.

R0 is estimated from the slope of the regression line:

$$
y = a + bx\\
dt = log(2)/b
$$


```{r}
# plot log graphs, doubling times for >100, last 14/30/60/90
log_x_start <- c(17,rows-14,rows-30,rows-60,rows-90)
log_title <- c("all dates","past 14 days","past 30 days","past 60 days","past 90 days")
n=0
dt = matrix(0,nrow=5,ncol = 3)
rownames(dt) = log_title
colnames(dt) = c("Argentina","CABA","Province of Buenos Aires")
for(i in log_x_start) {
  n=n+1
  for(val in x) {
    max_y <- floor(max(log(covid_frame[i:rows,locality["avg",val]])))
    plot(as.numeric(rownames(covid_frame[i:rows,])),
         log(covid_frame[i:rows,locality["avg",val]]),
         type="p",
         xlab="Days since 100 cases", 
         ylab="7-day rolling average (log)", 
         main=paste("New cases (log scale),",
                    locality["title",val],
                    "-",
                    log_title[n]
              ), 
         col=locality["light",val])
    minor.tick(nx=2,ny=2)
    grid(nx=0, ny=NULL,col="lightgray",lty="dotted",lwd=par("lwd"),equilogs=TRUE)
    fit <- lm(log(covid_frame[i:rows,locality["avg",val]]) 
              ~ as.numeric(rownames(covid_frame[i:rows,])))
    abline(fit, col=locality["dark",val])
    dt[log_title[n],locality["title",val]] <- round(log(2)/fit$coef[2],2) # Doubling time
    legend(x="topleft",
           y=1,
           adj=c(0,0),
           legend=c(paste("Doubling Time = ",
                          dt[log_title[n],
                             locality["title",val]],"days"))
    )
    
  }
}
dt
```

## R0 history

These graphs rely heavily on the Epitrix, EpiEstim, and incidence modules in R.

The following data on serial incidence are drawn from a meta analysis of COVID-19: https://doi.org/10.1002/jmv.26041

$$
\mu = 5.08\ days\\
\sigma = .18
$$
A gamma distribution is created programatically, and the estimate_R function is run against incidence objects containing the new cases reported each day.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# predict R0
# coronavirus details from https://doi.org/10.1002/jmv.26041
mu = 5.08  # mean incubation period
sigma = .18 # deviation in meta analysis
cv = mu/sigma
params <- gamma_mucv2shapescale(mu,cv)

incid_x_start <- c(0,21,37,67) # ends on 7-day rolling window
incid_title <- c("overall","past 14 days","past 30 days","past 60 days")
n=0

for(i in incid_x_start) {
  n=n+1
  for(val in x) {
    # hardcode at 8 to avoid rolling 7-day NAs
    incid <- as.incidence(covid_frame[8:rows,locality["avg",val]], 
                          dates=covid_frame[8:rows,1], 
                          interval=1)
    incid_rows <- dim(incid)
    end_x <- as.numeric(incid_rows[1])
    if(i==0) {
      start_x <- 1
    }
    else {
      start_x <- end_x-i
    }
    si <- distcrete("gamma", 
                  shape=params$shape, 
                  scale=params$scale, 
                  interval=1, w=0.5)
    #pred_1 <- project(incid, R=1.15, si=si, n_days=30, n_sim=1000)
    #f <- fit(incid[start_x:end_x])
    #plot (incid[start_x:end_x], fit=f, color=locality["light",val])
    #f
    res_parametric_si <- estimate_R(incid, 
                          method = "parametric_si",
                          config = make_config(list(
                          mean_si=mu,
                          std_si=sigma)))
    
    #@TODO get dates in the x axis labels
    #plot(res_parametric_si, legend=FALSE)
    plot(res_parametric_si$R[start_x:end_x,3],
         ylab="R0",
         xlab="Days",
         type="l",
         main=paste("R0 over time,",locality["title",val],incid_title[n]),
         col=locality["dark",val]
        )
    minor.tick(ny=2)
    #axis(1,incid$date,format(incid$date, "%b %d"), cex.axis = .7)
    grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
           lwd = par("lwd"), equilogs = TRUE)
    }
}
```

