---
title: "DS Flags Report"
author: "andino"
date: "02/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(incidence)
library(projections)
library(epitrix)
library(EpiModel)
library(EpiEstim)
library(rmarkdown)
library(knitr)
library(distcrete)
library(Hmisc)
library(tidyverse)
library(zoo)
library(reactable)
library(webshot)
library(htmlwidgets)
library(lubridate)
library(maps)
library(mapdata)
#library(data.table)


opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = '!h')
options(knitr.kable.NA = '')
webshot::install_phantomjs()

## GLOBAL FUNCTIONS
# Remove "Cases" and "Deaths" from headers
simplify_headers <- function(x) {
  colnames(x) <- gsub("Cases","",colnames(x))
  colnames(x) <- gsub("Deaths","",colnames(x))
  colnames(x) <- tolower(colnames(x))
  return(x)
}

# Remove common Spanish accents to allow string matching
remove_accents <- function(x) {
  x <- gsub("é","e",x)
  x <- gsub("á","a",x)
  x <- gsub("í","i",x)
  x <- gsub("ó","o",x)
  x <- gsub("ú","u",x)
  x <- gsub("É","E",x)
  x <- gsub("Á","A",x)
  x <- gsub("Í","I",x)
  x <- gsub("Ó","O",x)
  x <- gsub("Ú","U",x)
  x <- gsub("ñ","n",x)
  x <- gsub("Ñ","N",x)
  return(x)
}

# Find the per-100k value of a vector
per_100k <- function(x,y,gender="a") {
  g_row <- 1
  if(tolower(gender)=="m"){ 
    g_row <- 2 
  }
  if(tolower(gender)=="f"){ g_row <- 3 }
  colnames.x <- colnames(x)
  colnames.y <- colnames(y)
  start_col <- grep("conurbano",colnames.x) + 1
  end_col <- ncol(x)
  z <- x[1]
  for(i in 2:ncol(x)) {
    z[i] <- round(100000/as.numeric(populations[g_row,grep(colnames.x[i],colnames.y,ignore.case = TRUE)])*x[i],2)
  }
  colnames(z) <- colnames.x
  con_total <- rowSums(y[,start_col:end_col])
  z$conurbano <- round(100000/con_total[g_row]*x$conurbano,2)
  z <- as.data.frame(z)
  return(z)
}

# Get rolling mean and add back in date column
roll_set <- function(x,days=7,dates=x$date){
  days <- 7
  sub <- x[,2:ncol(x)]
  z <- round(rollmean(sub,days),2)
  z <- as.data.frame(cbind("date"=dates[days:nrow(sub)],z))
  return(z)
}

# Test whether strings match irrespective of accents, spacing, capitalization
string_match <- function(string1,string2) {
  string1 <- remove_accents(tolower(gsub(" ","",string1)))
  string2 <- remove_accents(tolower(gsub(" ","",string2)))
  test <- string1==string2
}

# Create column combining AMBA and ex-PBA
# @TODO error handling?
add_amba <- function(argdf) {
  argdf$amba <- argdf$conurbano + argdf$caba
  argdf$pbaexamba <- argdf$buenosaires - argdf$conurbano
  return(argdf)
}

## END GLOBAL FUNCTIONS

# VARIABLES
AMBA_start <- "2020-08-14"
roll_default <- 7

# Load data from scraped files

cases_file <- "data/cases.csv"
conurbano_file <- "data/conurbano.csv"
deaths_file <- "data/deaths.csv"
deaths_deltas_file <- "data/deaths_deltas.csv"
cases_deltas_file <- "data/cases_deltas.csv"
conurbano_deltas_file <- "data/conurbano_deltas.csv"
tests_file <- "data/tests.csv"
icu_rates_file <- "data/icu_rates.csv"
tests_deltas_file <- "data/test_deltas.csv"

conurbano_cum_data <- read.table(conurbano_file,header=TRUE,sep=",")
conurbano_deltas <- read.table(conurbano_deltas_file,header=TRUE,sep=",")
cases_cum_data <- read.table(cases_file,header=TRUE,sep=",")
cases_deltas <- read.table(cases_deltas_file,header=TRUE,sep=",")
death_cum_data <- read.table(deaths_file,header=TRUE,sep=",")
death_deltas <- read.table(deaths_deltas_file,header=TRUE,sep=",")
tests_data <- read.table(tests_file,header=TRUE,sep=",")
icu_rates_data <- read.table(icu_rates_file,header=TRUE,sep=",")

# Clean and combine case data
conurbano_dates <- conurbano_cum_data$Date
conurbano_delta_dates <-conurbano_deltas$Date
conurbano_cum_data <- conurbano_cum_data[-1]
conurbano_deltas <- conurbano_deltas[-1]
conurbano_cases_cum_data <- cbind("Date"=conurbano_dates,conurbano_cum_data[,seq(1,length(conurbano_cum_data),7)])
conurbano_deaths_cum_data <- cbind("Date"=conurbano_dates,conurbano_cum_data[,seq(2,length(conurbano_cum_data),7)])
conurbano_cases_deltas <- cbind("Date"=conurbano_delta_dates,conurbano_deltas[,seq(1,length(conurbano_deltas),7)])
conurbano_deaths_deltas <- cbind("Date"=conurbano_delta_dates,conurbano_deltas[,seq(2,length(conurbano_deltas),7)])

cumulative_cases <- merge(cases_cum_data,conurbano_cases_cum_data,by.x="Date",all.x=TRUE)
case_deltas <- merge(cases_deltas,conurbano_cases_deltas,by.x="Date",all.x=TRUE)
cumulative_deaths <- merge(death_cum_data,conurbano_deaths_cum_data,by.x="Date",all.x=TRUE)
death_deltas <- merge(death_deltas,conurbano_deaths_deltas,by.x="Date",all.x=TRUE)
cumulative_cases[is.na(cumulative_cases)] <- 0
case_deltas[is.na(case_deltas)] <- 0
cumulative_deaths[is.na(cumulative_deaths)] <- 0
death_deltas[is.na(death_deltas)] <- 0

#
# Redistribute death data
# PBA, in particular, underreported deaths from May-October.  Need to make a decision on how to smooth this.  
# For now, what we're doing is proportionally redistributing all of the deaths reported on October 1. 
# Oct 1 deaths set to average of week before and after it for each locality.
#
death_dates <- death_deltas$Date
death_adj_deltas <- death_deltas[-1]
death_adj_cum <- cumulative_deaths[-1]

for(i in 1:ncol(death_adj_deltas)) {
  oct1_deaths_series <- death_adj_deltas[41:55,i]
  oct1_deaths_series <- oct1_deaths_series[-8]
  oct1_deaths <- round(sum(oct1_deaths_series)/length(oct1_deaths_series))
  redistribute <- death_adj_deltas[48,i] - oct1_deaths
  locality_sum <- sum(death_adj_deltas[1:47,i])
  
  if(redistribute > 0) {
    for(n in 1:47) {
      death_adj_deltas[n,i] <- death_adj_deltas[n,i] + round(death_adj_deltas[n,i] / locality_sum * redistribute)
      #death_adj_cum[n,i] <- round(death_adj_cum[n,i] + death_adj_cum[n,i]/redistribute*death_adj_cum[n,i])
    }
  }
  # correct rounding errors by summing on October 1
  death_adj_deltas[48,i] <- oct1_deaths + sum(death_deltas[1:47,n+1]) - sum(death_adj_deltas[1:47,n])
}
death_adj_deltas <- cbind("Date"=death_dates,death_adj_deltas)
#death_adj_cum <- cbind("Date"=death_dates,death_adj_cum)


# Simplify column headers
base_data <- list("Cases"=cumulative_cases,
                  "CaseDeltas"=case_deltas,
                  "Deaths"=cumulative_deaths,
                  "DeathDeltas"=death_adj_deltas)
base_data <- lapply(base_data,simplify_headers)
dates <- cumulative_cases$Date
num_dates <- nrow(cumulative_cases)
dates.5 <- dates[5:num_dates]
dates.7 <- dates[7:num_dates]

# Add AMBA columns
base_data <- lapply(base_data,add_amba)

# Obtain rolling averages
base_data.5 <- lapply(base_data,roll_set,days=5)
base_data.7 <- lapply(base_data,roll_set,days=7)

# Obtain per 100k numbers
national_populations_file <- "data/PopulationNational.csv"
conurbano_populations_file <- "data/PopulationPBA.csv"

national_populations_t <- read.table(national_populations_file,header=TRUE,sep=",")
national_populations_t$TotalPopulation <- as.numeric(gsub(",","",national_populations_t$TotalPopulation))
national_populations_t$MalePopulation <- as.numeric(gsub(",","",national_populations_t$MalePopulation))
national_populations_t$FemalePopulation <- as.numeric(gsub(",","",national_populations_t$FemalePopulation))

national_populations_t[nrow(national_populations_t)+1,] <- c("National",sum(national_populations_t$TotalPopulation),
                                        sum(national_populations_t$MalePopulation),
                                        sum(national_populations_t$FemalePopulation))
national_populations <- data.frame(t(national_populations_t[-1])) 
colnames(national_populations) <- gsub(" ","",remove_accents(national_populations_t[,1]))
conurbano_populations_t <- read.table(conurbano_populations_file,header=TRUE,sep=",")
conurbano_populations_t$Municipality <- trimws(conurbano_populations_t$Municipality)
conurbano_populations_t$TotalPopulation <- as.numeric(gsub(",","",conurbano_populations_t$TotalPopulation))
conurbano_populations_t$MalePopulation <- as.numeric(gsub(",","",conurbano_populations_t$MalePopulation))
conurbano_populations_t$FemalePopulation <- as.numeric(gsub(",","",conurbano_populations_t$FemalePopulation))

conurbano_populations <- data.frame(t(conurbano_populations_t[-1]))
colnames(conurbano_populations) <- gsub(" ","",remove_accents(conurbano_populations_t$Municipality))
populations <- cbind(national_populations,conurbano_populations)

base_data.100k <- lapply(base_data,per_100k,y=populations)
base_data.5.100k <- lapply(base_data.100k,roll_set,5)
base_data.7.100k <- lapply(base_data.100k,roll_set,7)

delta_frame <- cbind(base_data$CaseDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=0,"per100k"=FALSE)
delta_frame <- rbind(delta_frame,
                     cbind(base_data.5$CaseDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=5,"per100k"=FALSE))
delta_frame <- rbind(delta_frame,
                     cbind(base_data.7$CaseDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=7,"per100k"=FALSE))
delta_frame <- rbind(delta_frame,
                     cbind(base_data.5.100k$CaseDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=5,"per100k"=TRUE))
delta_frame <- rbind(delta_frame,
                     cbind(base_data.7.100k$CaseDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=7,"per100k"=TRUE))
delta_frame <- rbind(delta_frame,
                     cbind(base_data.100k$CaseDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=0,"per100k"=TRUE))
delta_frame$date <- as.Date(delta_frame$date)
delta_frame$value <- as.numeric(delta_frame$value)

death_frame <- cbind(base_data$DeathDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=0,"per100k"=FALSE)
death_frame <- rbind(death_frame,
                     cbind(base_data.5$DeathDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=5,"per100k"=FALSE))
death_frame <- rbind(death_frame,
                     cbind(base_data.7$DeathDeltas %>% pivot_longer(!date,names_to="location", values_to="value"),
                     "roll"=7,"per100k"=FALSE))
death_frame$date <- as.Date(death_frame$date)
death_frame$value <- as.numeric(death_frame$value)

colnames(icu_rates_data) <- cbind("date","beds","national","amba")
icu_frame <- icu_rates_data %>% pivot_longer(!date,names_to="variable",values_to="value")
icu_frame$date <- as.Date(icu_frame$date)

provinces <- c("national"="National","caba"="CABA","buenosaires"="Province of Buenos Aires", 
               "amba"="AMBA (Conurbano + CABA)","catamarca"="Catamarca",
               "chaco"="Chaco","chubut"="Chubut","cordoba"="Córdoba","corrientes"="Corrientes",
               "entrerios"="Entre Ríos", "formosa"="Formosa","jujuy"="Jujuy",
               "lapampa"="La Pampa","larioja"="La Rioja","mendoza"="Mendoza",
               "misiones"="Misiones","neuquen"="Neuquén","rionegro"="Río Negro",
               "salta"="Salta","sanjuan"="San Juan","sanluis"="San Luis","santacruz"="Santa Cruz",
               "santafe"="Santa Fe","santiagodelestero"="Santiago del Estero",
               "tierradelfuego"="Tierra del Fuego","tucuman"="Tucumán")

amba_municipalites <- c("Almirante Brown","Avellaneda","Berazategui","Berisso","Brandsen",
          "Campana","Cañuelas","Ensenada","Escobar","Esteban Echeverría",
          "Exaltación de la Cruz","Ezeiza","Florencio Varela","General Las Heras",
          "General Rodríguez","General San Martín","Hurlingham","Ituzaingó",
          "José C. Paz","La Matanza","La Plata","Lomas de Zamora",
          "Luján","Marcos Paz","Malvinas Argentinas","Moreno","Merlo",
          "Morón","Pilar","Presidente Perón","Quilmes","San Fernando",
          "San Isidro","San Miguel","San Vicente","Tigre","Tres de Febrero",
          "Vicente López","Zárate")

BA_four <- c("national","amba","caba","buenosaires")


```

## New Cases

The following graphs show the overall epidemiological curves in the localities based on simple "new cases per day" as reported.  Note that date of case report DOES NOT equal date of first symptoms or diagnosis, necessarily, as there are frequently reporting delays.  Rather, this data is the change in cases from the previous day's Ministry of Health report:

```{r, total_cases, fig.show="hold",out.width="50%"}
for(this_province in BA_four) {
  ##NEW CODE
  if(string_match("amba",this_province)) {
    start_date <- as.Date(AMBA_start)
    first_month <- floor_date(as.Date(AMBA_start),"month")
  } else {
    start_date <- as.Date("2020-03-02")
    first_month <- as.Date("2020-03-01")
  }
  plot_frame <- delta_frame %>% filter(roll==0 & date >= start_date & per100k==FALSE & location == this_province)
  roll_frame <- delta_frame %>% filter(roll==roll_default & 
                                         date >= start_date & per100k==FALSE & location == this_province)
  
  peak_data <- top_n(roll_frame,n=1,wt=value)
  last_data <- roll_frame %>% filter(date==max(date))
  last_month <- ceiling_date(plot_frame$date[nrow(plot_frame)],"month")
  
   plot <- ggplot(plot_frame) +
    geom_col(plot_frame,mapping = aes(x=date,y=value,color="gray")) + 
    geom_line(roll_frame,mapping = aes(x=date,y=value,color="red")) +
    scale_x_continuous(name="Date",breaks=c(seq(from=first_month,to=last_month,by="month"))) +  
    scale_y_continuous(name="New Cases") +
    scale_color_manual(name="Series",labels=c("Cases","Rolling Average"),values=c("gray","red")) + 
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1), 
          legend.position="bottom",
          text=element_text(size=18)) + 
    labs(title=paste("New Cases, ",provinces[this_province],", ",roll_default,"-day rolling average",sep=""))  

  
  if(peak_data$value > last_data$value & (last_data$date - 7) > peak_data$date) {
    percent_change <- round((last_data$value - peak_data$value) / last_data$value * 100,2)
    plot <- plot + annotate("text",x=peak_data$date,y=peak_data$value,
                                             label=paste(peak_data$date,
                                                         round(peak_data$value),sep="\n"),
                                                         vjust=-.5,color="red")
    plot <- plot + annotate("text",x=last_data$date,y=last_data$value,
                          label=paste(round(last_data$value),
                                      "\n-",percent_change,"%",
                                      sep=""),vjust=1.5,color="red")
  
  } else {

   plot <- plot + annotate("text",x=last_data$date,y=last_data$value,
                          label=round(last_data$value),vjust=1.5,color="red")
  }
  print(plot)
  
}

```
\newpage

## 14-day trend

Phase 1: 14-day trend lines

```{r, two-week-trend, fig.show="hold",out.width="50%"}

start_date <- last_data$date - 14

for(this_province in BA_four) {
  
  plot_frame <- delta_frame %>% filter(roll==0 & date >= start_date & per100k==FALSE & location == this_province)
  roll_frame <- delta_frame %>% filter(roll==roll_default & 
                                         date >= start_date & per100k==FALSE & location == this_province)
  first_data <- roll_frame %>% filter(date==start_date)
  last_data <- roll_frame %>% filter(date==max(date))
  
  plot <- ggplot(plot_frame) +
    geom_col(plot_frame,mapping = aes(x=date,y=value,color="gray")) + 
    geom_line(roll_frame,mapping = aes(x=date,y=value,color="red")) +
    scale_x_continuous(name="Date",breaks=c(seq(from=first_month,to=last_month,by="day"))) +  
    scale_y_continuous(name="New Cases") +
    scale_color_manual(name="Series",labels=c("Cases","Rolling Average"),values=c("gray","red")) + 
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1), 
          legend.position="bottom",
          text=element_text(size=18)) + 
    labs(title=paste("New Cases past 14 days, ",provinces[this_province],
                     ", ",roll_default,"-day rolling average",sep="")) + 
    annotate("text",x=last_data$date,y=last_data$value,
                              label=round(last_data$value),vjust=1.5,color="red") +
    annotate("text",x=first_data$date,y=first_data$value,
                              label=round(first_data$value),vjust=-.5,color="red")
  
  print(plot)
}

```

\newpage

## 28-day trend

Phase 2: 28-day trend lines

```{r, four-week-trend, fig.show="hold",out.width="50%"}

start_date <- last_data$date - 28

for(this_province in BA_four) {
  
  plot_frame <- delta_frame %>% filter(roll==0 & date >= start_date & per100k==FALSE & location == this_province)
  roll_frame <- delta_frame %>% filter(roll==roll_default & 
                                         date >= start_date & per100k==FALSE & location == this_province)
  first_data <- roll_frame %>% filter(date==start_date)
  last_data <- roll_frame %>% filter(date==max(date))
  
  plot <- ggplot(plot_frame) +
    geom_col(plot_frame,mapping = aes(x=date,y=value,color="gray")) + 
    geom_line(roll_frame,mapping = aes(x=date,y=value,color="red")) +
    scale_x_continuous(name="Date",breaks=c(seq(from=first_month,to=last_month,by="week"))) +  
    scale_y_continuous(name="New Cases") +
    scale_color_manual(name="Series",labels=c("Cases","Rolling Average"),values=c("gray","red")) + 
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1), 
          legend.position="bottom",
          text=element_text(size=18)) + 
    labs(title=paste("New Cases past 28 days, ",provinces[this_province],
                     ", ",roll_default,"-day rolling average",sep="")) + 
    annotate("text",x=last_data$date,y=last_data$value,
                              label=round(last_data$value),vjust=1.5,color="red") +
    annotate("text",x=first_data$date,y=first_data$value,
                              label=round(first_data$value),vjust=-.5,color="red")
  
  print(plot)

}
```
\newpage
## ICU Capacity

```{r icu-graphs, message=FALSE, warning=FALSE}

first_icu_month <- floor_date(icu_frame$date[1],"month")
last_icu_month <- ceiling_date(icu_frame$date[nrow(icu_frame)],"month")
national_icu <- icu_frame %>% filter(variable=="national")
amba_icu <- icu_frame %>% filter(variable=="amba")
icu_plot <- ggplot(icu_frame,aes(x = date, na.rm=TRUE)) +   
    geom_line(data=national_icu,mapping=aes(x=date,y=value,na.rm=TRUE,color="blue"),size=1) + 
    geom_line(data=amba_icu,mapping=aes(x=date,y=value,na.rm=TRUE,color="red"),size=1) +
    scale_x_continuous(name="Date",breaks=c(seq(first_icu_month,last_icu_month, by="month"))) +
    scale_y_continuous(name="ICU occupancy percentage") +
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1),legend.position = "right") +
    scale_color_manual(name="Series",labels=c("National","AMBA"),values=c("blue","red")) +
    guides(fill = "legend") +
    labs(title="ICU Occupancy Rates")

print(icu_plot)
```

\newpage
## Testing status
Note:  This data is drawn from the Ministry of Health's determinaciones.csv file.  There are discrepancies in this data and the daily reports that MOH issues.

```{r positivity-graphs, message=FALSE,echo=FALSE}
tests_data <- read.table(tests_file,header=TRUE,sep=",")
tests_deltas_data <- read.table(tests_deltas_file,header=TRUE,sep=",")

colnames(tests_data) <- tolower(colnames(tests_data))
colnames(tests_deltas_data) <- tolower(colnames(tests_deltas_data))

tests_dates <- tests_data$date
tests_data <- tests_data[,-1]

end_loop <- ncol(tests_data)/2
for(i in 1:end_loop) {
  pos_col <- i*2
  total_col <- pos_col-1
  col_name <- paste("posratetests",gsub("totaltests","",colnames(tests_data[total_col])),sep="")
  tests_data[,col_name] <- tests_data[,pos_col] / tests_data[,total_col]
}
tests_data$date <- tests_dates

tests_frame <- tests_data %>% pivot_longer(!date,names_to=c("variable","location"),
                                             names_pattern="(.*)tests(.*)",
                                             values_to="tests",
                                             values_transform=list(tests=as.numeric))
tests_deltas_frame <- tests_deltas_data %>% pivot_longer(!date,names_to=c("variable","location"),
                                                          names_pattern="(.*)tests(.*)",
                                                          values_to="tests",
                                                          values_transform=list(tests=as.numeric))
tests_deltas_frame$date <- as.Date(tests_deltas_frame$date)
tests_frame$date <- as.Date(tests_frame$date)
tests_frame$location <- sub("^$","national",tests_frame$location)
tests_deltas_frame$location <- sub("^$","national",tests_deltas_frame$location)

tests_pos_frame <- filter(tests_frame,variable=="posrate")
first_tests_month <- floor_date(tests_frame$date[1],"month")
last_tests_month <- ceiling_date(tests_frame$date[nrow(tests_frame)],"month")


tests_pos_plot <- ggplot(tests_pos_frame,aes(x = date)) +   
    geom_line(data=tests_pos_frame %>% filter(location=="national"),
              mapping=aes(x=date,y=tests,na.rm=TRUE,color="blue"),size=1) +  
    geom_line(data=tests_pos_frame %>% filter(location=="caba"),
              mapping=aes(x=date,y=tests,na.rm=TRUE,color="yellow"),size=1) +
    geom_line(data=tests_pos_frame %>% filter(location=="buenosaires"),
              mapping=aes(x=date,y=tests,na.rm=TRUE,color="red"),size=1) +
    scale_x_continuous(name="Date",breaks=c(seq(first_tests_month,last_tests_month, by="month"))) +
    scale_y_continuous(name="Tests Positivity Rate") +
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1),legend.position = "right") +
    scale_color_manual(name="Series",labels=c("National","CABA","Province of Buenos Aires"),
                       values=c("blue","yellow","red")) +
    guides(fill = "legend") +
    labs(title="Test Positivitiy Rate")

## @TODO actually do test deltas.  Oops.
tests_plot <- ggplot(tests_deltas_frame,aes(x = date)) +   
    geom_line(data=tests_deltas_frame %>% filter(location=="national" & variable=="total"),
              mapping=aes(x=date,y=tests,na.rm=TRUE,color="blue"),size=1) +  
    geom_line(data=tests_deltas_frame %>% filter(location=="caba" & variable=="total"),
              mapping=aes(x=date,y=tests,na.rm=TRUE,color="yellow"),size=1) +
    geom_line(data=tests_deltas_frame %>% filter(location=="buenosaires" & variable=="total"),
              mapping=aes(x=date,y=tests,na.rm=TRUE,color="red"),size=1) +
    scale_x_continuous(name="Date",breaks=c(seq(first_tests_month,last_tests_month, by="month"))) +
    scale_y_continuous(name="Tests Conducted Per Day") +
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1),legend.position = "right") +
    scale_color_manual(name="Series",labels=c("National","CABA","Province of Buenos Aires"),
                       values=c("blue","yellow","red")) +
    guides(fill = "legend") +
    labs(title="Tests Conducted Per Day")

print(tests_pos_plot)
print(tests_plot)


```

\newpage
## Doubling times
The regression line is drawn using the R "lm()" function over the x values.

R0 is estimated from the slope of the regression line:

$$
y = a + bx
$$

$$
dt = log(2)/b
$$
```{r doubling-time-graphs, message=FALSE, warning=FALSE}

for(this_province in BA_four) {
  
  if(string_match("amba",this_province)) {
    start_date <- as.Date(AMBA_start)
    first_month <- floor_date(as.Date(AMBA_start),"month")
  } else {
    start_date <- as.Date("2020-03-02")
    first_month <- as.Date("2020-03-01")
  }

  plot_frame <- delta_frame %>% filter(roll==0 & date >= start_date & per100k==FALSE & location == this_province)
  plot_frame$value <- na_if(plot_frame$value,0)
  plot_frame$log <- log(plot_frame$value)
  plot_frame$day <- seq.int(nrow(plot_frame))
  
  first_data <- roll_frame %>% filter(date==start_date)
  last_data <- roll_frame %>% filter(date==max(date))
  fit <- lm(plot_frame$log ~ plot_frame$day) 
  dt <- round(log(2)/fit$coef[2],2) # Doubling time
  
  plot <- ggplot(plot_frame,aes(plot_frame$date,plot_frame$log,col="blue",shape=".")) +
    stat_summary(fun.plot_frame=mean_cl_normal) +
    #stat_smooth(method="lm",formula = plot_frame$log ~ plot_frame$day,color="red") +
    stat_smooth(method="lm", col="red") +
    scale_y_continuous(name="Daily cases (log)") +
    scale_x_continuous(name="Date",breaks=c(seq(from=first_month,to=last_month,by="month"))) +
    scale_color_manual(name="Series",labels=c("Cases","Line of Best Fit"),values=c("blue","red")) +
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1),
          legend.position="bottom") +
    labs(title=paste("Doubling Time, ",provinces[this_province],
                     ": ",dt," days",sep="")) 

  
  print(plot)

}

```
\newpage
## Better R Estimate

This data is drawn from over 1 million epidemiological records, indexed by the date the case was registered with the Ministry of Health. Cases are often registered prior to a confirmed diagnosis; therefore, this data "lags".  The dates are taken based on the date a case was "opened," rather than "confirmed" or "reported".  

An incidence object is created using all confirmed cases in Argentina. The estimate_R() function from the EpiEstim package is used with the serial interval described below. While the estimate_R() function uses a rolling 7-day window, we also cut off the final five days of data due to reduce the confirmation lag.

The following data on serial incidence are drawn from a meta analysis of COVID-19: https://doi.org/10.1002/jmv.26041

$$
\mu = 5.08\ days
$$

$$
\sigma = .18
$$
A gamma distribution is created programatically, and the estimate_R function is run against incidence objects containing the new cases reported each day.

```{r label=better-r-estimate, message=FALSE, warning=FALSE, fig.show="hold",out.width="50%"}
## Other indicators
load("CovidEpiFile.RData")

mu = 5.08  # mean incubation period
sigma = .18 # deviation in meta analysis
cv = mu/sigma
params <- gamma_mucv2shapescale(mu,cv)

confirmed <- raw_data[clasificacion_resumen %in% c("Confirmado")]
caba_incid <- confirmed[residencia_provincia_nombre %in% c("CABA")]
pba_incid <- confirmed[residencia_provincia_nombre %in% c("Buenos Aires")]
conurbano_incid <-pba_incid[residencia_departamento_nombre %in% amba_municipalites]
amba_incid <- rbind(caba_incid,conurbano_incid)
#sanisidro_incid <- pba_incid[residencia_departamento_nombre %in% c("San Isidro")]
loop_pools <- list(confirmed,caba_incid,amba_incid,conurbano_incid)
graph_titles <- c("National","CABA","AMBA","Conurbano")
graph_colors <- c("blue","green","red","red")

first_month <- as.Date("2020-04-01")
last_month <- ceiling_date(as.Date(today()),"month")

for(i in 1:length(loop_pools)) {
  incid <- incidence(as.Date(loop_pools[[i]]$fecha_apertura))
  right_bound <- nrow(incid$counts) - 5
  thirty <- right_bound - 37
  
  res_parametric_si <- estimate_R(incid[1:right_bound],
                                  method = "parametric_si",
                                  config = make_config(list(
                                        mean_si=mu,
                                        std_si=sigma)))
  plot_frame <- data.frame("date"=tail(head(res_parametric_si$dates,-5),-2) + 7,
                          "mean_r"=res_parametric_si$R[,3],
                          "lower_r"=res_parametric_si$R[,6],
                          "upper_r"=res_parametric_si$R[,10])
  plot_frame <- plot_frame %>% filter(date>="2020-04-01")
  
  plot_r <- ggplot(plot_frame,aes(x=plot_frame$date, y=plot_frame$mean_r,color="R0")) +
    geom_line() +
    geom_ribbon(aes(ymin=plot_frame$lower_r,ymax=plot_frame$upper_r),linetype=2, alpha=0.1) +
    scale_y_continuous(name="Rate of Transmission (95% CI)") +
    scale_x_continuous(name="Date",breaks=c(seq(from=first_month,to=last_month,by="month"))) +
    scale_color_manual(name="Series",labels=c("R0"),values=c("blue")) +
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1),
          legend.position="bottom") +
    labs(title=paste("Rate of Transmission, ",graph_titles[i],sep="")) 
  print(plot_r)
  
  plot_frame_60 <- tail(plot_frame,60)
  
  plot_r_60 <- ggplot(plot_frame_60,aes(x=plot_frame_60$date, y=plot_frame_60$mean_r,color="R0")) +
  geom_line() +
  geom_ribbon(aes(ymin=plot_frame_60$lower_r,ymax=plot_frame_60$upper_r),linetype=2, alpha=0.1) +
  scale_y_continuous(name="Rate of Transmission (95% CI)") +
  scale_x_continuous(name="Date",breaks=c(seq(from=floor_date(plot_frame_60$date[1],"week"),
                                              to=ceiling_date(plot_frame_60$date[60],"week"),by="week"))) +
  scale_color_manual(name="Series",labels=c("R0"),values=c("blue")) +
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1),
        legend.position="bottom") +
  labs(title=paste("Rate of Transmission, ",graph_titles[i]," past 60 days",sep="")) 
  print(plot_r_60)
}
```
\newpage
## New Cases (per 100k)

```{r label=provinces_per_100k, message=FALSE, warning=FALSE,fig.show="hold",out.width="33%"}

for(this_province in names(provinces)) {
  if(string_match("amba",this_province)) {
    start_date <- as.Date(AMBA_start)
    first_month <- floor_date(as.Date(AMBA_start),"month")
  } else {
    start_date <- as.Date("2020-03-02")
    first_month <- as.Date("2020-03-01")
  }
  last_month <- ceiling_date(today(),"month")
  
  plot_frame <- delta_frame %>% filter(roll==roll_default & date >= start_date & per100k==TRUE & location == this_province)
  last_data <- plot_frame[nrow(plot_frame),]
  
  per100k_plot <- ggplot(plot_frame) +   
    geom_line(data=plot_frame,mapping=aes(x=date,y=value,na.rm=TRUE,color="blue"),size=1) +  
    scale_x_continuous(name="Date",breaks=c(seq(first_month,last_month, by="quarter"))) +
    scale_y_continuous(name="Daily Cases per 100k residents") +
    theme(legend.position="none") +
    scale_color_manual(values=c("blue")) +
    annotate("text",x=last_data$date,y=last_data$value,
                              label=round(last_data$value),vjust=1.5,color="red") +
    labs(title=paste(provinces[this_province],": ",round(last_data$value)," daily cases per 100k"))
  
  print(per100k_plot)
}

```

\newpage
## Deaths
Note:  On October 1, the government reported nearly 3,000 deaths that had actually occurred between May and September.  The graphs below smooth the one-day anomaly by assigning those deaths proportionally over the 6 weeks running August 15-September 30. 

```{r label=deaths, message=FALSE, warning=FALSE,fig.show="hold",out.width="50%"}

for(this_province in BA_four) {
  ##NEW CODE

  start_date <- as.Date(AMBA_start)
  first_month <- floor_date(as.Date(AMBA_start),"month")

  plot_frame <- death_frame %>% filter(roll==0 & date >= start_date & per100k==FALSE & location == this_province)
  roll_frame <- death_frame %>% filter(roll==roll_default & 
                                         date >= start_date & per100k==FALSE & location == this_province)
  
  peak_data <- top_n(roll_frame,n=1,wt=value)
  last_data <- roll_frame %>% filter(date==max(date))
  last_month <- ceiling_date(plot_frame$date[nrow(plot_frame)],"month")
  
   plot <- ggplot(plot_frame) +
    geom_col(plot_frame,mapping = aes(x=date,y=value,color="gray")) + 
    geom_line(roll_frame,mapping = aes(x=date,y=value,color="red")) +
    scale_x_continuous(name="Date",breaks=c(seq(from=first_month,to=last_month,by="month"))) +  
    scale_y_continuous(name="Deaths") +
    scale_color_manual(name="Series",labels=c("Deaths","Rolling Average"),values=c("gray","red")) + 
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1), 
          legend.position="bottom",
          text=element_text(size=18)) + 
    labs(title=paste("Deaths, ",provinces[this_province],", ",roll_default,"-day rolling average",sep=""))  

  
  if(peak_data$value > last_data$value & (last_data$date - 7) > peak_data$date) {
    percent_change <- round((last_data$value - peak_data$value) / last_data$value * 100,2)
    plot <- plot + annotate("text",x=peak_data$date,y=peak_data$value,
                                             label=paste(peak_data$date,
                                                         round(peak_data$value),sep="\n"),
                                                         vjust=-.5,color="red")

    plot <- plot + annotate("text",x=last_data$date,y=last_data$value,label=paste(round(last_data$value),
                                       "\n-",percent_change,"%",
                                       sep=""),vjust=1.5,color="red")
  } else {

   plot <- plot + annotate("text",x=last_data$date,y=last_data$value,
                          label=round(last_data$value),vjust=1.5,color="red")
  }
  print(plot)
  
}
```