---
title: "Argentina Covid Report"
author: "Chris Andino"
date: "October 14 2020"
output: pdf_document
---
Data as of 9:30 am 15-OCT-2020

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(incidence)
library (projections)
library(epitrix)
library(EpiModel)
library(EpiEstim)
library(rmarkdown)
library(knitr)
library(distcrete)
library(Hmisc)
require(RCurl)
require(googlesheets4)
require(tidyr)
library(zoo)
library(reactable)
library(webshot)
library(htmlwidgets)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = '!h')
options(knitr.kable.NA = '')
webshot::install_phantomjs()

local_file <- "data/cases.csv"
conurbano_file <- "data/pba.csv"
conurbano_data <- read.csv(conurbano_file,header=TRUE,sep=",")
conurbano_dates <- conurbano_data$Date
conurbano_data$Date <- NULL
conurbano_data <- conurbano_data[,c(TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)]
conurbano_data$ConurbanoTotal <- rowSums(conurbano_data)
conurbano_data$Date <- conurbano_dates

data <- read.csv(local_file,header=TRUE,sep=",")
data <- merge(x=data, y=conurbano_data,by="Date",all.x=TRUE)
rows <- nrow(data)
NewCasesNational = c(0)
NewCasesCABA = c(0)
NewCasesPBA = c(0)
NewCasesAMBA = c(0)
CasesAMBA = c(data[1,"CasesCABA"]+data[1,"ConurbanoTotal"])
for(i in 2:rows) {
  NewCasesNational <- c(NewCasesNational,data[i,"CasesNational"]-data[i-1,"CasesNational"])
  NewCasesCABA <- c(NewCasesCABA,data[i,"CasesCABA"]-data[i-1,"CasesCABA"])
  NewCasesPBA <- c(NewCasesPBA,data[i,"CasesBuenosAires"]-data[i-1,"CasesBuenosAires"])
  CasesAMBA <- c(CasesAMBA,data[i,"CasesCABA"]+data[i,"ConurbanoTotal"])
  NewCasesAMBA <- c(NewCasesAMBA,CasesAMBA[i]-CasesAMBA[i-1])
}
AvgCasesNational = c(0,0,0,0,round(rollmean(NewCasesNational,5)))
AvgCasesCABA = c(0,0,0,0,round(rollmean(NewCasesCABA,5)))
AvgCasesPBA = c(0,0,0,0,round(rollmean(NewCasesPBA,5)))
AvgCasesAMBA = c(0,0,0,0,round(rollapply(NewCasesAMBA,width=5,mean,na.rm=TRUE)))
AvgCasesAMBA <- (na.locf(AvgCasesAMBA) + rev(na.locf(rev(AvgCasesAMBA))))/2 # fill NAs
covid_frame <- data.frame(Date=data$Date,
                          TotalCasesNational=data$CasesNational,
                          NewCasesNational,
                          AvgCasesNational,
                          TotalCasesCABA=data$CasesCABA,
                          NewCasesCABA,
                          AvgCasesCABA,
                          TotalCasesPBA=data$CasesBuenosAires,
                          NewCasesPBA,
                          AvgCasesPBA,
                          TotalCasesAMBA=CasesAMBA,
                          NewCasesAMBA,
                          AvgCasesAMBA)

covid_frame[,1] <- as.Date(covid_frame[,1],"%Y-%m-%d")

# Set up for loops
national <- c("Argentina","NewCasesNational","AvgCasesNational","light blue","blue")
CABA <- c("CABA","NewCasesCABA","AvgCasesCABA","green","dark green")
PBA <- c("Conurbano","NewCasesPBA","AvgCasesPBA","pink","red")
AMBA <- c("AMBA","NewCasesAMBA","AvgCasesAMBA","yellow","dark orange")
locality = array(c(national,CABA,PBA,AMBA),
                 dim=c(5,4)
                 )
colnames(locality) <- c("Argentina","CABA","PBA","AMBA (CABA+Conurbano)")
rownames(locality) <- c("title","cases","avg","light","dark")
x <- colnames(locality)

```

## New Cases

The following graphs show the overall epidemiological curves in the localities based on simple "new cases per day" as reported.  Note that date of case report DOES NOT equal date of first symptoms or diagnosis, necessarily.  Rather, this data is the change in cases from the previous day's report:

```{r label=new-cases, message=FALSE, warning=FALSE}
# plot total cases
for(val in x) {
  plot(covid_frame[1:rows,1],
       covid_frame[1:rows,locality["cases",val]],
       type="n",
       xlab="Date", 
       ylab="Cases",
       main=paste("Daily new cases, ",locality["title",val])
  )
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  minor.tick(nx=2,ny=2)
  lines(covid_frame[1:rows,1],
        covid_frame[1:rows,locality["cases",val]],
        col=locality["light",val])
  lines(covid_frame[1:rows,1],
        covid_frame[1:rows,locality["avg",val]],
        col=locality["dark",val])
  legend(x="topleft",
         y=max(covid_frame[1:rows,locality["cases",val]],na.rm=T),
         legend=c("Daily Report","5-Day Rolling Average"),
         col=c(locality["light",val],locality["dark",val]),
         lty=1:2, cex=0.8)
}
start_frame = rows-14
show_frame <- as.data.frame(covid_frame[start_frame:rows,])
show_frame
#rtable <- reactable(show_frame,minRows=15)
#html <- "rtable.html"
#saveWidget(rtable,html)
#webshot(html, "rtableSnapshot.png")
```
\newpage

## 14-day trend

Phase 1: 14-day trend lines

```{r 14-day-trend, message=FALSE, warning=FALSE}
# plot 14-day trend
  start_x = rows-14
  #rows = rows-2
for(val in x) {

  plot(covid_frame[start_x:rows,1],
       covid_frame[start_x:rows,locality["cases",val]],
       type="n",
       xlab="Date", 
       ylab="Cases",
       main=paste("14-day trend, ",locality["title",val])
  )
  #minor.tick(ny=2)
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  points(covid_frame[start_x:rows,1],
        covid_frame[start_x:rows,locality["cases",val]],
        col=locality["light",val])
  lines(covid_frame[start_x:rows,1],
        covid_frame[start_x:rows,locality["avg",val]],
        col=locality["dark",val])
  #fit <- lm(covid_frame[start_x:rows,locality["avg",val]] ~ covid_frame[start_x:rows,1])
  #abline(fit)
  legend(x="topleft",
         y=max(covid_frame[1:rows,locality["cases",val]],na.rm=T),
         legend=c("Daily Report","5-Day Rolling Average"),
         col=c(locality["light",val],locality["dark",val]),
         lty=1:2, cex=0.8)
}
```
## Phase 2 decisions

```{r 28-day-trend, message=FALSE, warning=FALSE}
# plot 28-day trend
  start_x = rows-28
  #rows = rows-2
for(val in x) {

  plot(covid_frame[start_x:rows,1],
       covid_frame[start_x:rows,locality["cases",val]],
       type="n",
       xlab="Date", 
       ylab="Cases",
       main=paste("28-day trend, ",locality["title",val])
  )
  #minor.tick(ny=2)
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  points(covid_frame[start_x:rows,1],
        covid_frame[start_x:rows,locality["cases",val]],
        col=locality["light",val])
  lines(covid_frame[start_x:rows,1],
        covid_frame[start_x:rows,locality["avg",val]],
        col=locality["dark",val])
  #fit <- lm(covid_frame[start_x:rows,locality["avg",val]] ~ covid_frame[start_x:rows,1])
  #abline(fit)
  legend(x="topleft",
         y=max(covid_frame[1:rows,locality["cases",val]],na.rm=T),
         legend=c("Daily Report","5-Day Rolling Average"),
         col=c(locality["light",val],locality["dark",val]),
         lty=1:2, cex=0.8)
}
```
\newpage

## Log graphs

The following graphs are generated by:
$$
x = Number\ of\ Days\ since\ March\ 3\
$$

$$
y = log(Number\ of\ New\ Cases\ this\ day)
$$
The regression line is drawn using the R "lm()" function over the x values.

R0 is estimated from the slope of the regression line:

$$
y = a + bx
$$
$$
dt = log(2)/b
$$


```{r log-graphs}
# plot log graphs, doubling times for >100, last 14/30/60/90
log_x_start <- c(17,rows-14,rows-30,rows-60,rows-90)
log_title <- c("all dates","past 14 days","past 30 days","past 60 days","past 90 days")
n=0
dt = matrix(0,nrow=5,ncol = 4)
rownames(dt) = log_title
colnames(dt) = c("Argentina","CABA","PBA") # remove AMBA+Conurbano because problems
for(i in log_x_start) {
  n=n+1
  for(val in x) {
    max_y <- floor(max(log(covid_frame[i:rows,locality["avg",val]])))
    plot(as.numeric(rownames(covid_frame[i:rows,])),
         log(covid_frame[i:rows,locality["avg",val]]),
         type="p",
         xlab="Days since 100 cases", 
         ylab="5-day rolling average (log)", 
         main=paste("New cases (log scale),",
                    locality["title",val],
                    "-",
                    log_title[n]
              ), 
         col=locality["light",val])
    minor.tick(nx=2,ny=2)
    grid(nx=0, ny=NULL,col="lightgray",lty="dotted",lwd=par("lwd"),equilogs=TRUE)
    fit <- lm(log(covid_frame[i:rows,locality["avg",val]]) 
              ~ as.numeric(rownames(covid_frame[i:rows,])))
    abline(fit, col=locality["dark",val])
    dt[log_title[n],locality["title",val]] <- round(log(2)/fit$coef[2],2) # Doubling time
    legend(x="topleft",
           y=1,
           adj=c(0,0),
           legend=c(paste("Doubling Time = ",
                          dt[log_title[n],
                             locality["title",val]],"days"))
    )
    
  }
}
dt
```
\newpage

## R0 over time (daily cases estimate)

These graphs rely heavily on the Epitrix, EpiEstim, and incidence modules in R. These graphs are rough estimates based on the number of new cases reported each day and not/not the actual date of registry/onset of symptoms, which provide a more-accurate picture of the rate of transmission.

The following data on serial incidence are drawn from a meta analysis of COVID-19: https://doi.org/10.1002/jmv.26041

$$
\mu = 5.08\ days
$$
$$
\sigma = .18
$$
A gamma distribution is created programatically, and the estimate_R function is run against incidence objects containing the new cases reported each day.

```{r weak-r-estimate, echo=FALSE, message=FALSE, warning=FALSE}
# # predict R0
# # coronavirus details from https://doi.org/10.1002/jmv.26041
# mu = 5.08  # mean incubation period
# sigma = .18 # deviation in meta analysis
# cv = mu/sigma
# params <- gamma_mucv2shapescale(mu,cv)
# 
# incid_x_start <- c(0,37)
# #incid_x_start <- c(0,21,37,67) # ends on 7-day rolling window
# #incid_title <- c("overall","past 14 days","past 30 days","past 60 days")
# incid_title <-c ("overall", "past 30 days")
# n=0
# 
# for(i in incid_x_start) {
#   n=n+1
#   for(val in x) {
#     # hardcode at 7 to avoid rolling 7-day NAs
#     incid <- as.incidence(covid_frame[7:rows,locality["avg",val]], 
#                           dates=covid_frame[7:rows,1], 
#                           interval=1)
#     incid_rows <- dim(incid)
#     end_x <- as.numeric(incid_rows[1])
#     if(i==0) {
#       start_x <- 1
#     }
#     else {
#       start_x <- end_x-i
#     }
#     # si <- distcrete("gamma", 
#     #               shape=params$shape, 
#     #               scale=params$scale, 
#     #               interval=1, w=0.5)
#     
#     #pred_1 <- project(incid, R=1.15, si=si, n_days=30, n_sim=1000)
#     #f <- fit(incid[start_x:end_x])
#     #plot (incid[start_x:end_x], fit=f, color=locality["light",val])
#     #f
#     res_parametric_si <- estimate_R(incid, 
#                           method = "parametric_si",
#                           config = make_config(list(
#                           mean_si=mu,
#                           std_si=sigma)))
#     
#     #@TODO get dates in the x axis labels
#     #plot(res_parametric_si, legend=FALSE)
#     plot(res_parametric_si$R[start_x:end_x,3],
#          ylab="R0",
#          xlab="Days",
#          type="l",
#          main=paste("R0 over time,",locality["title",val],incid_title[n]),
#          col=locality["dark",val]
#         )
#     minor.tick(ny=2)
#     #axis(1,incid$date,format(incid$date, "%b %d"), cex.axis = .7)
#     grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
#            lwd = par("lwd"), equilogs = TRUE)
#     }
# }
```



```{r deaths, message=FALSE, warning=FALSE}
## Other indicators

### Deaths
```
\newpage
### ICU Capacity


```{r icu-graphs, message=FALSE, warning=FALSE}
  icu_file <- "data/icu_rates.csv"
  icu_data <- read.csv2(icu_file,sep=",")
  plot(as.Date(icu_data$Date),
       icu_data$ICUPctAMBA,
       ylim = c(0,100),
       type="n",
       xlab="Date", 
       ylab="Cases",
       main=paste("Daily ICU Bed Rate")
  )
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  #minor.tick(nx=2,ny=2)
  points(as.Date(icu_data$Date),
        icu_data$ICUPctNation,
        col="blue")
  points(as.Date(icu_data$Date),
        icu_data$ICUPctAMBA,
        col="red")
  legend(x="topleft",
         y=max(icu_data$ICUPctAMBA,na.rm=T),
         legend=c("Nationwide","AMBA"),
         col=c("blue", "red"),
         lty=1:2, cex=0.8)

tail(icu_data,30)
```

## Better R Estimate

This data is drawn from over 1 million epidemiological records, indexed by the date the case was registered with the Ministry of Health. Cases are often registered prior to a confirmed diagnosis; therefore, this data "lags".  

An incidence object is created using all confirmed cases in Argentina. The estimate_R() function from the EpiEstim package is used with the serial interval as described in the R estimate section above. While the estimate_R() function uses a rolling 7-day window, we also force the estimate away from the last five days of data due to the confirmation lag.

```{r label=better-r-estimate, message=FALSE, warning=FALSE }
## Other indicators
load("CovidEpiFile.RData")

mu = 5.08  # mean incubation period
sigma = .18 # deviation in meta analysis
cv = mu/sigma
params <- gamma_mucv2shapescale(mu,cv)

confirmed <- raw_data[clasificacion_resumen %in% c("Confirmado")]
caba_incid <- confirmed[residencia_provincia_nombre %in% c("CABA")]
pba_incid <- confirmed[residencia_provincia_nombre %in% c("Buenos Aires")]
amba_incid <- rbind(caba_incid,pba_incid)
sanisidro_incid <- pba_incid[residencia_departamento_nombre %in% c("San Isidro")]
loop_pools <- list(confirmed,caba_incid,pba_incid,sanisidro_incid)
graph_titles <- c("National","CABA","Conurbano","AMBA","San Isidro")
graph_colors <- c("blue","green","orange","red","red")

for(i in 1:length(loop_pools)) {
  incid <- incidence(as.Date(loop_pools[[i]]$fecha_apertura))
  right_bound <- nrow(incid$counts) - 5
  thirty <- right_bound - 37
  
  res_parametric_si <- estimate_R(incid[1:right_bound],
                                  method = "parametric_si",
                                  config = make_config(list(
                                        mean_si=mu,
                                        std_si=sigma)))
  
  plot(res_parametric_si$R[,3],
       ylab="R0",
       xlab="Days",
       type="l",
       main=paste("R0 over time,",graph_titles[i],"Overall"),
       col=graph_colors[i],
       ylim=c(.5,2.5)
  )
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
  
  plot(res_parametric_si$R[thirty:right_bound,3],
       ylab="R0",
       xlab="Days",
       type="l",
       main=paste("R0 over time,",graph_titles[i],"Past Month"),
       col=graph_colors[i],
       ylim=c(.5,2)
  )
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
}


```

## Testing and positivity rates

```{r positivity_rates, message=FALSE, warning=FALSE}
test_data <- read.csv2("data/tests.csv",header=TRUE,sep=",")
NewTestsNational = c(test_data$TotalTests[1])
PositiveTestsNational = c(test_data$PositiveTests[1])
PositivityNational = c(PositiveTestsNational/NewTestsNational)
NewTestsCABA = c(test_data$TotalTestsCABA[1])
PositiveTestsCABA = c(test_data$PositiveTestsCABA[1])
PositivityCABA = c(PositiveTestsCABA/NewTestsCABA)
NewTestsPBA = c(test_data$TotalTestsBuenosAires[1])
PositiveTestsPBA = c(test_data$PositiveTestsBuenosAires[1])
PositivityPBA = c(PositiveTestsPBA/NewTestsPBA)

test_rows = nrow(test_data)
for(i in 2:test_rows) {
  NewTestsNational = c(NewTestsNational,test_data$TotalTests[i]-test_data$TotalTests[i-1])
  PositiveTestsNational = c(PositiveTestsNational,test_data$PositiveTests[i]-test_data$PositiveTests[i-1])
  PositivityNational = c(PositivityNational,PositiveTestsNational[i]/NewTestsNational[i])
  NewTestsCABA = c(NewTestsCABA,test_data$TotalTestsCABA[i]-test_data$TotalTestsCABA[i-1])
  PositiveTestsCABA = c(PositiveTestsCABA,test_data$PositiveTestsCABA[i]-test_data$PositiveTestsCABA[i-1])
  PositivityCABA = c(PositivityCABA,PositiveTestsCABA[i]/NewTestsCABA[i])
  NewTestsPBA = c(NewTestsPBA,test_data$TotalTestsBuenosAires[i]-test_data$TotalTestsBuenosAires[i-1])
  PositiveTestsPBA = c(PositiveTestsPBA,test_data$PositiveTestsBuenosAires[i]-test_data$PositiveTestsBuenosAires[i-1])
  PositivityPBA = c(PositivityPBA,PositiveTestsPBA[i]/NewTestsPBA[i])
  nrow(PositiveTestsNational)
}
test_df <- as.data.frame(cbind(as.Date(test_data$Date),NewTestsNational,PositiveTestsNational,PositivityNational,
                           NewTestsCABA,PositiveTestsCABA,PositivityCABA,NewTestsPBA,PositiveTestsPBA,
                           PositivityPBA),col.names=TRUE)
test_locality = c("National","CABA","Province of Buenos Aires")
for(i in 1:3) {
  y_data = i*3+1
  plot(test_df$V1,test_df[,y_data],
       main=paste("Test positivity rate",test_locality[i]),
       xlim=c(test_df$V1[1],test_df$V1[test_rows]),
       xlab="Date",
       ylab="Positivity rate",
       ylim=c(0,1))
  grid(nx = 0, ny = NULL, col = "lightgray", lty = "dotted",
       lwd = par("lwd"), equilogs = TRUE)
}
test_df
```